---
title: "CPSC541_BYOD_Project"
author: "Yuting_Qiu"
date: "9/19/2019"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read in the data and basic processing
```{r}
data1 <-read.csv("Moatt et al Data_S1.csv",h=T)
data2 <- data1[data1$Include == "y",] #only inlude =y
data2$Diet <- as.factor(data2$Diet)
data3 <- data2[-which(is.na(data2$Pprotein )),] # remove the nas
```


```{r}
str(data3)
```

```{r}
data4 <- data3[,c(4,5,6,7,8,9,12,13,16,17,19,21,25)] # the possible columns that we are interested in
pairs(data4) # making scatterplot matrices of the data 
```

```{r}
# fit some basic model to see the results
# effect on lipid intake on body protein:lipid intake

M_lipid <- lm(P_F ~ L, data4)
M_protein <- lm(P_F ~ P, data4)
M_PLratio <- lm(P_F ~ P_L, data4)


plot(data4$L, data4$P_F)
lines(data4$L, predict(M_lipid),col="red")

plot(data4$P, data4$P_F)
lines(data4$P, predict(M_protein),col="red")

plot(data4$P_L, data4$P_F)
lines(data4$P_L, predict(M_PLratio), col="red")

summary(M_lipid)
summary(M_protein)
summary(M_PLratio)


```

evaluate the diet on body composion 
```{r}
data3$P <- as.numeric(as.character(data3$P))
data3$L <- as.numeric(as.character(data3$L))
data3$E <- as.numeric(as.character(data3$E))
data3$P_F <- as.numeric(as.character(data3$P_F))

m.diet1 <- lm(P_F ~ Diet, data = data3)
m.diet2 <- lm(P_F ~ P + L + E + P_L, data = data3)
m.diet3 <- lm(P_F ~  P_L, data = data3)
m.diet4 <- lm(P_F ~  P_L + L, data = data3)
m.diet5 <- lm(P_F ~  P_L + P, data = data3)
m.diet6 <- lm(P_F ~  P_L + P + L, data = data3)
m.diet7 <- lm(P_F ~  Diet + Sex, data = data3)
m.diet8 <- lm(P_F ~  Diet + Sex + Tank, data = data3)
m.diet9 <- lm(P_F ~  Diet + Tank, data = data3)
m.diet9 <- lm(P_F ~  E, data = data3)
m.diet10 <- lm(P_F ~  E + P_L, data = data3)
m.diet11 <- lm(E ~  P_L, data = data3)
m.diet12 <- lm(P_F ~  L, data = data3)
m.diet13 <- lm(P_F ~  P, data = data3)


anova(m.diet1)
anova(m.diet2)
anova(m.diet3)
anova(m.diet4)
anova(m.diet5)
anova(m.diet6)
anova(m.diet9)
anova(m.diet10)
anova(m.diet11)

summary(m.diet1)
summary(m.diet2)
summary(m.diet3)
summary(m.diet4)
summary(m.diet5)
summary(m.diet6)
summary(m.diet7) # Adjusted R-squared:  0.3428 
summary(m.diet9)
summary(m.diet11)
summary(m.diet12)
summary(m.diet13)
summary(lm(E~L, data = data3)) # Adjusted R-squared:  0.9465 
summary(lm(P_F~L, data = data3)) # Adjusted R-squared:  0.3293
summary(lm(P_F~P, data = data3)) # Adjusted R-squared:  -0.007666 
summary(lm(P_F~E, data = data3)) # Adjusted R-squared:  0.3377 
summary(lm(P_F~P_L, data = data3)) # Adjusted R-squared:  0.2577 



anova(m.diet3, m.diet4)
anova(m.diet3, m.diet4)
anova(m.diet3, m.diet6) # it is important both P and L amonunt in the diet and the P_L ratio
anova(m.diet1, m.diet7) # beside diet, sex is also an important facotor
anova(m.diet7, m.diet8) # tank is not signigficant with sex and diet
anova(m.diet1, m.diet9) # tank is not signficant

# test for collinearity in the P, L and P_L
summary(lm(P_F ~ P + L + P_L, data = data3))
summary(lm(P_F ~  L + P_L + P, data = data3))
summary(lm(P_F ~ P_L + P + L , data = data3))
```



# re-fit, try poisson errors
```{r}
m_lipid.2 <- glm(P_F ~ L, data = data4, family = poisson(link = identity)) # non-interger, cannot run poission
```

# try poly-nominal
```{r}
m_lipid.3 <- lm(P_F ~ poly(L,2), data4)
plot(data4$L, data4$P_F, xlab = "lipid content in the diet", ylab = "Protein: fat ratio in the fish body", main = "Polynomial model")
lines(data4$L, predict(m_lipid.3),col="red")
summary(m_lipid.3)

m_lipid.4 <- lm(P_F ~ poly(L,3), data4)
plot(data4$L, data4$P_F, xlab = "lipid content in the diet", ylab = "Protein: fat ratio in the fish body", main = "Polynomial model")
lines(data4$L, predict(m_lipid.4),col="red")
summary(m_lipid.4)

anova(m_lipid.3, m_lipid.4)
# not significant, p is large, complicate model is not necessary better than the simple model
anova(m_lipid.3, M_lipid)
# not signifciant, a simple model is good enough
# but we can still see the influence of lipid intake on the protein_ratio for the fish. more lipid intake, less protein:fat ratio, not much influence for the protein
```

# look at the body componient 
```{r}
data5 <- data3[,12:25]
str(data5)

pairs(data5)


```


```{r}
# what can explain protein:fat ration in the sample
# fat percentage first, looks like ploy normal

library(ggplot2)
p1 = ggplot(data5, aes(x=Pfat, y=P_F))
p1 = p1 + geom_point() + ylab("protein:fat ratio in the sample") +
  xlab("percentage of fat in the sample")
p1
```

```{r}
# fit polynomial to the P:L sample and 
fit1 <- lm(P_F~poly(Pfat,2), data5)
plot(fit1)
summary(fit1)
```

CHECK FOR OUTLIERS
```{r}
influence.measures(fit1)
```


```{r}
plot(data5$Pfat, data5$P_F, xlab = "percentage of fat in the sample", ylab = "protein: fat ratio in fish body", main = "Fit with polynomial distribution")
lines(data5$Pfat, predict(fit1), col="red")
```

since tank is a significant factor, we need to regress it out
```{r}
model_tank <- lm(data3$P_F ~ data3$Tank)
anova(model_tank)
tank_resid = residuals(model_tank)
data3$tank_resid<- tank_resid
plot(model_tank)

```




```{r}
# fit the exponential model to the data
fit_exp <- glm(P_F ~ Pfat, data = data5, family = gaussian(link=log))
plot(fit_exp)

# check residuals
# res_exp <- residuals(fit_exp)/sqrt(summary(fit_exp)$dispersion) 

# ggplot(res_exp, aes(x=predict(fit_exp), y=res_exp)) + geom_point()
```


 ############################### 
 after the first feedback
 ##############################


```{r}
# esclude the unwanted columns
data_ready <- data3[,-c(1:3,5:8,11,22:24,26:31)]

# compare tank and family, tank is significant
model_blocking <- lm(data3$P_F ~ data3$Tank + data3$Family)
anova(model_blocking)
model_blocking <- lm(data3$P_F ~  data3$Family + data3$Tank)
anova(model_blocking)

# how anout sex
model_blocking <- lm(data3$P_F ~  data3$Sex + data3$Tank)
anova(model_blocking)
# both sex and tank is significant, family is not significant and is excluded in the prior code

anova(lm(data3$P_F ~ data3$Pfat + data3$Tank))
anova(lm(data3$P_F ~ data3$Pfat + data3$Sex))
```



```{r}

# try to conduct stepwise regression
data_transformed <- na.omit(data_ready)
model_null <- lm(P_F ~ 1, data=data_transformed)
model_full <- lm(P_F ~ ., data = data_transformed)
step(model_null, scope = list(upper = formula(model_full), lower = formula(model_null)), direction = 'forward')
step(model_full, scope = list(upper = formula(model_full), lower = formula(model_null)), direction = 'backward')
step(model_full, scope = list(upper = formula(model_full), lower = formula(model_null)), direction = 'both')

#### FORWARD
# Coefficients:
# (Intercept)         Pfat          Fat  Fat_free_wt     Pprotein         Sexm  
#    6.48586     -0.22609     22.37837     -7.92262      0.03346     -0.19282  


#### BACKWARD
# Coefficients:
# (Intercept)       Wet_wt  Fat_free_wt     Ashed_wt      Protein       Pwater         Pfat         Pash  
# -5.537e-16    2.498e-01   -2.212e+00    3.723e-01    1.660e+00   -7.440e-02   -1.180e+00   -1.105e-01    

#### BOTH
# Coefficients:
# (Intercept)       Wet_wt       Dry_wt     Ashed_wt      Protein       Pwater         Pash     Pprotein  
#    -3.5310       8.2526      12.9655     -35.9563     -57.7010      -0.2718       0.2532       0.3545 

#########
# why is tank not show up here????
# why P_L not showing up????



```
 

since there is no significant value of outliers may I can use the original dataset
```{r}
anova(lm(P_F~Sex, data = data3))
anova(lm(P_F~Wet_wt, data = data3))
anova(lm(P_F~Dry_wt, data = data3))
anova(lm(P_F~Ashed_wt, data = data3))
anova(lm(P_F~Fat, data = data3))
anova(lm(P_F~Protein, data = data3))
anova(lm(P_F~Pfat, data = data3))
anova(lm(P_F~Pprotein, data = data3))
anova(lm(P_F~Energy_fat, data = data3))
anova(lm(P_F~Energy_protein, data = data3))
anova(lm(P_F~Total_energy, data = data3))
# all are significant

summary(lm(P_F~Sex, data = data3)) # Adjusted R-squared:  0.03802 
summary(lm(P_F~Wet_wt, data = data3)) # Adjusted R-squared:  0.224 
summary(lm(P_F~Dry_wt, data = data3)) # Adjusted R-squared:  0.2973
summary(lm(P_F~Ashed_wt, data = data3)) # Adjusted R-squared:  0.04432 
summary(lm(P_F~Protein + Fat, data = data3)) # Adjusted R-squared:  0.4895
summary(lm(P_F~Pfat, data = data3)) #Adjusted R-squared:  0.7854 
summary(lm(P_F~Pprotein, data = data3)) # Adjusted R-squared:  0.5719 
summary(lm(P_F~Energy_fat, data = data3)) # Adjusted R-squared:  0.4464 
summary(lm(P_F~Energy_protein, data = data3)) # Multiple R-squared:  0.2076
summary(lm(P_F~Total_energy, data = data3)) # Adjusted R-squared:  0.3664 

cor.test(data3$Wet_wt, data3$Dry_wt)

data_ready <- data3[,-c(1:3,5:8,11,22:24,26:31)]

# since both sex and tank is significant, we need to regress them out first
# then scale the residuals for log regression 

model.tank_sex <- lm(data_ready$P_F ~ data_ready$Sex + data_ready$Tank)
anova(model.tank_sex)
resid(model.tank_sex) -> data_ready$resid

library(caret)
data_stand <- preProcess(data_ready, method = c("center", "scale", "YeoJohnson","nzv"))
data_transformed <- predict(data_stand, data_ready)
rm.out<-function(x)(ifelse(x >3, NA,x ))

dev.new()
pairs(data_ready)

model.5 <- lm(resid ~ poly(Pfat,2), data_ready)
plot(data_ready$Pfat, data_ready$resid)
lines(data_ready$Pfat, predict(model.5),col="red")
summary(model.5)

#fit_exp <- glm(data_ready$resid ~ Pfat, data = data_ready, family = gaussian(link=log))
#plot(fit_exp)
# regress out the factor is not going to work, with residuals and the map look to be much more noiser

# how about look femanle and male fish seperately?
library(dplyr)
data_female <- filter(data_ready, Sex =="f")
data_male <- filter(data_ready, Sex =="m")

p <- ggplot(data_ready,aes(Pfat,P_F,colour = Sex)) + 
  geom_point() +
  ylab("Protein: fat ratio in the fish body") +
  xlab("Percentage of fat in the fish body") +
  ggtitle("Distribution of female and male fish")

p


fit_exp.f <- glm(P_F ~ Pfat, data = data_female, family = gaussian(link=log))
plot(fit_exp.f)

fit_exp.m <- glm(P_F ~ Pfat, data = data_male, family = gaussian(link=log))
plot(fit_exp.m)

model.5 <- lm(resid ~ poly(Pfat,2), data_ready)
plot(data_ready$Pfat, data_ready$resid)


library(gridExtra)
library(ggplot2)
library(GGally)
# pairs plot to visualize data by sex (3 categories)

sexplot<-ggpairs(data_ready[,-3], aes(colour=Sex, alpha=0.8), title ="Pairs plot for fish dataset by sex")+
  theme_grey(base_size = 8)
pdf("sexplot.pdf")
sexplot
dev.off()
```

# try to adjust for tank
```{r}
grand_mean <- mean(data_ready[["P_F"]])
grand_mean #[1] 2.62429

# extract mean for each tank and then correct it
tank_level <- levels(data3$Tank)
tank_mean <- NULL

as.data.frame(aggregate(data_ready[,14], list(data_ready$Tank),mean)) -> tank_mean
names(tank_mean) <- c("tankid", "tank_mean")
tank_mean$grand_mean <- 2.62429
tank_mean$diff <- tank_mean$tank_mean - tank_mean$grand_mean

data_correction <- data_ready[,-15]


# goal is to put diff of tank_mean into the correlation datafrmae
data_correction <- merge(x=data_correction, y=tank_mean[,c(1,4)], by.x = "Tank", by.y = "tankid")
data_correction$new_pf <- data_correction$P_F - data_correction$diff # new_pf is the one thay has been corrected

# plot it after correction
library(ggplot2)
p <- ggplot(data_correction,aes(Pfat,new_pf,colour = Sex)) + geom_point()
p
# it is a distastet

# it seems that p_fat = 16.1, p_diff = 3.03 is an outlier it is data point 90, need to removeit 
data_correction <- data_correction[-90,]
p <- ggplot(data_correction,aes(Pfat,P_F,colour = Sex)) + geom_point() + geom_text(aes(label=rownames(data_correction)),hjust=0, vjust=0)
p
# now I am able to detect the two outliers, are they truely outliers????

# try differnt fit and compare them
data_f <- data_ready[data_ready$Sex=="f",]
data_m <- data_ready[data_ready$Sex=="m",]

# fit log link to it
fit_exp <- glm(P_F ~ Pfat, data = data_correction, family = gaussian(link=log))
plot(fit_exp)
fit_exp_f <- glm(P_F ~ Pfat, data = data_f, family = gaussian(link=log))
plot(fit_exp_f)

fit_exp_m <- glm(P_F ~ Pfat, data = data_m, family = gaussian(link=log))
plot(fit_exp_m)

# poly-nomial
fit_poly <- lm(P_F~poly(Pfat,2), data_correction)
plot(fit_poly)
summary(fit_poly)

fit_poly_f <- lm(P_F~poly(Pfat,2), data_correction[data_correction$Sex == "f", ])
plot(fit_poly_f)
summary(fit_poly_f)

fit_poly_m <- lm(P_F~poly(Pfat,2), data_correction[data_correction$Sex == "m", ])
plot(fit_poly_m)
summary(fit_poly_m)

# inverse.gamma
fit_gamma <- glm(P_F ~ Pfat, data = data_correction, family = Gamma(link = "identity"))
plot(fit_gamma)

fit_inverse <- glm(P_F ~ Pfat, data = data_correction, family = Gamma(link = "inverse"))
plot(fit_inverse)

# poission
fit_log_poi <- glm(P_F ~ Pfat, data = data_correction, family = poisson(link = log))
plot(fit_log_poi)

fit_log_poi_f <- glm(P_F ~ Pfat, data = data_correction[data_correction$Sex == "f",], family = poisson(link = log))
plot(fit_log_poi_f)

fit_log_poi_m <- glm(P_F ~ Pfat, data = data_correction[data_correction$Sex == "m",], family = poisson(link = log))
plot(fit_log_poi_m)

AIC(fit_exp) #[1] 77.66344  this seem to be a good model
AIC(fit_exp_f) # 91.33888
AIC(fit_exp_m) # -8.987106
AIC(fit_log_poi) #[1] Inf
AIC(fit_inverse) #[1] -54.14454 
AIC(fit_poly) # [1] 97.01369
AIC(fit_gamma) #[1] 140.0517           
```


```{r}
# fit_exp is the model that we are going to use
fit_exp <- glm(P_F ~ Pfat, data = data_correction, family = gaussian(link=log))
data_correction$predict_fitexp <- predict(fit_exp, type = "response")
exp_plot <-  ggplot(data = data_correction,aes(x=P_F, y=predict_fitexp)) + geom_point() +
  geom_smooth() +
  xlab("Observed protein: fat ratio in the fish body") +
  ylab("Predicted protein: fat ratio in the fish body") +
  ggtitle("Predicted v.s observed")
exp_plot

```


## it is possible to predict the P_F ratio of a fish based on their diet
```{r}
library(corrplot)
corr <- cor(data_correction[,3:14])
corrplot(corr, method="circle")

pairs(data_correction[,3:14])
## that would be a logisrtic regression later
```


## simulation test
## Simulation
Please simulate a set of 1000 new observations based on your best model.
How do these simulated observations compare to the real observations?

```{r}
# objective: new y based on the model fit, we need intercept and 5 parameters and also error 
# need a dataframe of x, and also random error ()

summary(data_correction) # mean, median and variance, min and max, mean and sd can give a normal distributioin (pnorm)

Pfat <- seq(min(data_correction$Pfat), max(data_correction$Pfat),length.out = 1000)
grid <- expand.grid(Pfat)
names(grid) <- c("Pfat")

summary(fit_exp)
# (Dispersion parameter for gaussian family taken to be 0.1130409)
# it is the same as mean square error

set.seed(777)
grid$error <- rnorm(n=1000, mean=0, sd=0.1130409)
head(grid)

library(tidyverse)
grid_sim <- grid %>%
  mutate(y1 = 2.514625 -0.065279*Pfat + error) %>%
  mutate(y2 = exp(y1))

grid_sim$pred <- predict(fit_exp, newdata = grid_sim) # based on the new dataframe of the dataframe, calculate the response without the error

plot(x=grid_sim$y1, y=grid_sim$pred, xlab = "simulated protein: fat ratio", ylab = "predicted protein: fat ratio")
cor.test(x = grid_sim$y1, y=grid_sim$pred, method = "pearson")
abline(a=0,b=1)
# this looks great
```


```{r}
# see if we are able to find PL ratio based on the diet the fish take
data6 <- data3[,c(4:12,25)]
data6 <- data3[,c(6:9,12,19,25)]
pairs(data6)

# based on the paired plot, it seems that overall, lipid is a very good indicator, and the inclusion of proteion would even mess up the correlation
library(corrplot)
corr <- cor(data6)
corrplot.mixed(corr)
corrplot(corr, method="circle")

```


